---
title: "How to continue the inference algorithm"
author: "Xavier Didelot, Jake Carson"
date: '`r Sys.Date()`'
output: html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{How to continue the inference algorithm}
  %\usepackage[utf8]{inputenc}
---
```{r echo=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4) 
```

If you want to reproduce exactly the same results as the ones shown in this tutorial, you need to set the seed of your random number generator to zero:
```{r}
library(TransPhylo2)
set.seed(0)
```

We simulate an outbreak that starts in 2005 and which and is observed up to 2008, across two locations:

```{r}
simu <- simulateOutbreak(kappa=0.25,lambda=0.25,pi=0.25,off.r=5,w.shape=10,
                         w.scale=0.1,dateStartOutbreak=2005,dateT=2008,ndemes=2,rho=0.8)
```

This simulation contains both the transmission tree between infected hosts and the within-host phylogenetic tree of each host. This can be visualised as a colored phlogenetic tree, where each host is represented by a unique color:

```{r}
plot(simu)
plotdemesCTree(simu)
```

The transmission tree can be extracted and plotted separately from the phylogeny:


```{r}
ttree=extractTTree(simu)
plot(ttree,type='detailed',w.shape=10,w.scale=0.1)
```

The phylogenetic tree can be extracted and plotted separately from the transmission tree:

```{r}
ptree<-extractPTree(simu)
p<-phyloFromPTree(ptree)
plot(p)
ape::axisPhylo(backward = F)
```

Set a seed and run the inference algorithm

```{r,results='hide'}
ptree<-ptreeFromPhylo(p,dateLastSample=dateLastSample(simu),demes=simu$demes)
set.seed(1)
res<-inferTTree(ptree,mcmcIterations=1000,w.shape=10,w.scale=0.1,dateT=2008,update.rho=T)
```

```{r}
plot(res)
```

We can replicate these results in multiple stages. First perform a shorter MCMC run:
```{r,results='hide'}
set.seed(1)
res_a<-inferTTree(ptree,mcmcIterations=500,w.shape=10,w.scale=0.1,dateT=2008,update.rho=T)
```

Then continue:
```{r,results='hide'}
res_b<-inferTTreecont(res_a,mcmcIterations=500)
```

Typically, the results should be consistent across either approach when using the same seed. However, small numerical differences can add up over repeated iterations. 
```{r}
chain1 <- extractParmTrace(res)
chain2 <- extractParmTrace(res_b)
plot(c(chain1[, 1]), type = "l", ylab = "off.r, one-stage appraoch")
plot(c(chain2[, 1]), type = "l", ylab = "off.r, two-stage appraoch")
```
