---
title: "Example where demes have different offspring distribution and sampling probabilities"
author: "Xavier Didelot, Jake Carson"
date: '`r Sys.Date()`'
output: html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example where demes have different offspring distribution and sampling probabilities}
  %\usepackage[utf8]{inputenc}
---
```{r echo=F}
knitr::opts_chunk$set(fig.width=6, fig.height=4) 
```

If you want to reproduce exactly the same results as the ones shown in this tutorial, you need to set the seed of your random number generator to zero:
```{r}
library(TransPhylo2)
set.seed(0)
```

We simulate an outbreak that starts in 2005 and which and is observed up to 2008, across two locations:

```{r}
simu <- simulateOutbreakmulti(kappa=0.25,lambda=0.25,pi=c(0.25, 0.75),off.r=c(2, 5),w.shape=10,
                         w.scale=0.1,dateStartOutbreak=2005,dateT=2008,ndemes=2,rho=c(0.6, 0.8))
```

This simulation contains both the transmission tree between infected hosts and the within-host phylogenetic tree of each host. This can be visualised as a colored phlogenetic tree, where each host is represented by a unique color:

```{r}
plot(simu)
plotdemesCTree(simu)
```

The transmission tree can be extracted and plotted separately from the phylogeny:


```{r}
ttree=extractTTree(simu)
plot(ttree,type='detailed',w.shape=10,w.scale=0.1)
```

The phylogenetic tree can be extracted and plotted separately from the transmission tree:

```{r}
ptree<-extractPTree(simu)
p<-phyloFromPTree(ptree)
plot(p)
ape::axisPhylo(backward = F)
```


```{r,results='hide'}
ptree<-ptreeFromPhylo(p,dateLastSample=dateLastSample(simu),demes=simu$demes)
res<-inferTTreemulti(ptree,mcmcIterations=1000,w.shape=10,w.scale=0.1,dateT=2008,update.rho=T)
```

```{r}
plot(res)
```
